%using document class from KOMA-Script
\documentclass{scrartcl}
\title{M1F Summary Notes}
\subtitle{JMC Year 1, 2017/2018 syllabus}
\date{}
\author{Fawaz Shah}

% Packages for adding hyperlinks to table of contents
\usepackage{color}   %May be necessary if you want to color links
\usepackage{hyperref}
\hypersetup{
    colorlinks=true, %set true if you want colored links
    linktoc=all,     %set to all if you want both sections and subsections linked
    linkcolor=black,  %choose some color if you want links to stand out
}

%package that allows aligned equations
\usepackage{amsmath}

%package that allows notation for extra mathematical symbols
\usepackage{amssymb}

%new commands for popular sets for ease of use
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}

% renaming command to writing vectors in bold notation
\renewcommand{\vec}[1]{\mathbf{#1}}

%package for managing images
\usepackage{graphicx}
\graphicspath{ {../img/} }

%package for managing hyperlinks
\usepackage{hyperref}

%package for added blue colored boxes
\usepackage[most]{tcolorbox}

%package that allowed removing indent from enumerate environment
\usepackage{enumitem}

%package that allows for negating nearly any symbol
\usepackage{centernot}

\begin{document}
\large
\maketitle
\noindent This document contains a list of definitions and a list of theorems.
\\\\
Note that the exam will probably require you to PROVE some of these theorems, so you should refer back to the original notes for the proofs.
\\\\
Boxes cover content in more detail. Titles of some theorems are given in italics.
\tableofcontents
\newpage

\section{Definitions}

\paragraph{Arbitrary union}
We can define the arbitrary union
\begin{equation}
\bigcup_{i \in I} X_{i}
\end{equation}
to be the union of all sets $ X_{i} $.

\paragraph{Arbitrary intersection}
We can define the arbitrary intersection
\begin{equation}
\bigcap_{i \in I} X_{i}
\end{equation}
to be the intersection of all sets $ X_{i} $.

\paragraph{Modulus (complex number)}
The modulus of $ z = x + iy $ is defined as:
\begin{equation}
|z| = \sqrt{x^{2} + y^{2}}
\end{equation}

\paragraph{Argument (complex number)}
The argument of $ z = x + iy $ is defined as:
\begin{equation}
\arg(z) = \arctan(\frac{y}{x})
\end{equation}

\paragraph{Fundamental Theorem of Algebra}
If a polynomial is of degree $ n $ then it has $ n $ roots in the complex plane.

\paragraph{Rational}
A number is rational if it can be expressed in the form $ \frac{p}{q} $, where $ p, q \in \Z $ and $ \gcd(p, q) = 1 $.

\paragraph{Root of unity}
Any $ z $ that is a solution to $ z^{n} = 1 $.

\paragraph{"Dense"}
A set $ S $ is dense in $ T $ if for any $ t_{1}, t_{2} \in T $ there exists an $ s \in S $ such that $ t_{1} < s < t_{2} $

\paragraph{Supremum}
A real number $ b $ is the sup (least upper bound) of a set $ S $ if:
\begin{itemize}
\item every $ s \in S $ is less than $ b $ (i.e. $ b $ is an upper bound)
\item $ b $ is less than every other upper bound of $ S $
\end{itemize}

\paragraph{Infimum}
A real number $ b $ is the inf (greatest lower bound) of a set $ S $ if:
\begin{itemize}
\item every $ s \in S $ is greater than $ b $ (i.e. $ b $ is a lower bound)
\item $ b $ is greater than every other lower bound of $ S $
\end{itemize}

\paragraph{Bounded above}
A set is bounded above if it has an upper bound.

\paragraph{Bounded below}
A set is bounded below if it has a lower bound.

\paragraph{Coprime}
We say two integers $ a $ and $ b $ are coprime if $ \gcd(a, b) = 1 $.

\paragraph{Fundamental Theorem of Arithmetic}
Every integer greater than 1 can be expressed as a unique product of primes.

\paragraph{$ \equiv$ (congruence operator)}
We say $ a \equiv b \mod m $ if $ (a - b) $ is a multiple of $ m $.

\paragraph{Binary relation}
A binary relation $ \sim $ on a set $ S $ is a subset of $ S \times S $.

\paragraph{Reflexive}
A binary relation $ \sim $ is reflexive if for any $ a \in S $:
\begin{equation}
a \sim a
\end{equation}

\paragraph{Symmetric}
A binary relation $ \sim $ is symmetric if:
\begin{equation}
a \sim b \Leftrightarrow b \sim a
\end{equation}

\paragraph{Transitive}
A binary relation $ \sim $ is transitive if:
\begin{equation}
a \sim b \land b \sim c \Rightarrow a \sim c
\end{equation}

\paragraph{Equivalence relation}
A binary relation $ \sim $ is an equivalence relation iff it is reflexive, symmetric and transitive.

\paragraph{Equivalence class}
The equivalence class of $ b \in S $ is the set of all $ s \in S $ such that $ b \sim s $.

\paragraph{Image}
The image of a function $ f: A \to B $ is the set of all $ b \in B $ such that $ b = f(a) $ for some $ a \in A $.

\paragraph{Injective}
A function $ f : A \to B $ is injective if every $ b \in B $ has at MOST one $ a \in A $ that maps to it. Mathematically:
\begin{equation}
\forall x, y \in A \quad f(x) = f(y) \Rightarrow x = y
\end{equation}

\paragraph{Surjective}
A function $ f : A \to B $ is surjective if every $ b \in B $ has at LEAST one $ a \in A $ that maps to it. Mathematically:
\begin{equation}
\forall b \in B \quad \exists a \in A \quad s.t. \quad f(a) = b
\end{equation}

\paragraph{Bijective}
A function $ f : A \to B $ is bijective iff it is both surjective and injective.

\paragraph{(Two-sided) inverse function}
The inverse of a function $ f : A \to B $ is defined as $ f^{-1} : B \to A $, and satisfies:
\begin{align}
f^{-1} \circ f \equiv f \circ f^{-1} \equiv Id_{A}
\end{align}

\paragraph{Power set}
The power set of $ S $ is the set of all subsets of $ S $.

\paragraph{Countably infinite}
An infinitely-sized set $ S $ is countably infinite if there exists a bijection $ f : \Z_{> 0} \to S $.

\section{Theorems}

\subsection{Logic and Sets}
\textit{Contrapositive law}
\\
For any Boolean statements $ P, Q $:
\begin{equation}
(P \Rightarrow Q) \Leftrightarrow (\neg Q \Rightarrow \neg P)
\end{equation}
\\
To negate any logical statements, we can flip the $ \forall $ and $ \exists $ signs and negate the predicate (the mathematical statement at the end).

\subsection{Complex numbers}
Complex numbers can be added, subtracted, multiplied and divided.
\\\\
For any complex number $ z $,
\begin{equation}
z \overline{z} = |z| ^{2}
\end{equation}
\\
For any complex numbers $ z_{1}, z_{2} $:
\begin{align}
\overline{z_{1} + z_{2}} & = \overline{z_{1}} + \overline{z_{2}} \\
\overline{z_{1} - z_{2}} & = \overline{z_{1}} - \overline{z_{2}} \\
\overline{z_{1} z_{2}} & = \overline{z_{1}} \cdot \overline{z_{2}} \\
\overline{(\frac{z_{1}}{z_{2}})} & = \frac{\overline{(z_{1})}}{\overline{(z_{2})}} \\
\overline{(z_{1})^{n}} & = (\overline{z_{1}})^{n}
\end{align}
The inverse of a complex number $ z $ is such that:
\begin{equation}
z z^{-1} = 1
\end{equation}
\\
Any complex number $ z $ can be represented in exponential form:
\begin{equation}
z = r e^{i \theta}
\end{equation}
where $ r $ is the modulus of $ z $ and $ \theta $ is the argument.
\\\\
\textit{Euler's formula}
\begin{equation}
e^{i \theta} = \cos(\theta) + i \sin(\theta)
\end{equation}
This can be proven using power series.
\\\\
\textit{De Moivre's Theorem}
\\
For any complex number $ z $:
\begin{equation}
z^{n} = r^{n} (\cos(n \theta) + i \sin(n \theta))
\end{equation}
This can be proven by induction or using Euler's formula.
\\\\
\textit{Fundamental Theorem of Algebra}
\\
Any polynomial of degree $ n $ has $ n $ roots in the complex plane.
\\\\
\textit{Roots of unity}
\\
Any root of unity (i.e. a solution to $ z^{n} = 1 $) can be expressed in the form:
\begin{equation}
z = e^{\frac{2 \pi k i}{n}}, \quad k \in \Z_{\geq 0}
\end{equation}
If we let:
\begin{equation}
\omega = e^{\frac{2 \pi i}{n}}
\end{equation}
then all the $ n^{th} $ roots of unity can be expressed in the form:
\begin{equation}
z = 1, \omega, \omega^{2}, \omega^{3} \textrm{ ... } \omega^{n-1}
\end{equation}
In general, all $ n^{th} $ roots of a complex number $ a $ can be expressed as:
\begin{equation}
z = a, a\omega, a\omega^{2}, a\omega^{3} \textrm{ ... } a\omega^{n-1}
\end{equation}

\subsection{Proofs and Induction}
\textit{Counterexample}
\\
This involves disproving a statement by providing an example situation where the statement is false.
\\\\
\textit{Proof by contradiction}
\\
Assume the statement is true and reach a contradiction. This means the statement must be false.
\\\\
\textit{Weak induction}
\\
If we take $ P $ as a predicate:
\begin{equation}
P(0) \land (P(k) \Rightarrow P(k+1)) \Rightarrow P(n) \quad (\forall n)
\end{equation}
Note that we do not have to start at 0:
\begin{equation}
P(m) \land (P(k) \Rightarrow P(k+1)) \Rightarrow P(n) \quad (\forall n \geq m)
\end{equation}
\\
\textit{Strong induction}
\\
If we take $ P $ as a predicate:
\begin{equation}
P(0) \land ((\forall j \leq k \ P(j)) \Rightarrow P(k+1)) \Rightarrow P(n) \quad (\forall n)
\end{equation}
Note that we do not have to start at 0:
\begin{equation}
P(m) \land ((\forall j \leq k \ P(j)) \Rightarrow P(k+1)) \Rightarrow P(n) \quad (\forall n \geq m)
\end{equation}
\begin{tcolorbox}[breakable]
We must prove $ P(k) \Rightarrow P(k+1) $ for weak induction.
\\
We must prove $ (\forall j \leq k \ P(j)) \Rightarrow P(k+1) $ for strong induction.
\\\\
Strong induction is mathematically equivalent to weak induction.
\end{tcolorbox}

\subsection{Number Theory}
There is no smallest positive real number.
\\\\
There always exists a real $ n^{th} $ root of a real number (for $ n \in Z_{>0} $).
\\\\
The sum/product of any two rational numbers is rational.
\\\\
The sum of a rational number and an irrational number is irrational.
\\
(proof by contradiction)
\\\\
There are infinitely many irrational numbers.
\\
(consider $ x_{n} = \sqrt{2} + n $ for $ n \in \Z $)
\\\\
$ \Z $ is unbounded in $ \R $ (i.e. for any real number there is always an integer larger than it).
\\\\
$ \Q $ is dense in $ \R $.
\\\\
$ \R \setminus \Q $ (i.e. the set of irrationals) is dense in $ \R $.
\\\\
\textit{Completeness Axiom}
\\
A set $ S $ has a least upper bound (sup) iff:
\begin{itemize}
\item $ S $ is non-empty
\item $ S $ is bounded above
\end{itemize}
Corresponding statement for greatest lower bound (inf).
\\\\
Every real number has a decimal expansion.
\\
\begin{tcolorbox}[breakable]
Assume for simplicity that we have a single digit real number.
\\
Let $ d_{n} $ be the sequence of digits and $ e_{n} $ the sequence of error terms, such that $ \forall n \ 0 \leq e_{n} \leq 1 $. We have:
\begin{equation}
x = d_{0}.d_{1}d_{2}d_{3}...
\end{equation}
Then we can produce the digits by the following steps:
\begin{align}
x & = d_{0} + e_{0} \\
10e_{0} & = d_{1} + e_{1} \\
10e_{1} & = d_{2} + e_{2} \\
& \vdots \\
10e_{n-1} & = d_{n} + e_{n}
\end{align}
\end{tcolorbox}
\noindent
\\
A number is rational $ \Leftrightarrow $ it has a periodic decimal expansion.
\\\\
Any integer $ a $ has a quotient and remainder when divided by an integer $ b $.
\\
(proof by induction on $ a $)
\\\\
This quotient and remainder are unique for any $ a $.
\\
(proof by contradiction)
\\\\
For any $ x \in \Z_{\geq 0} $ and $ \lambda, \mu \in \Z,  (x \ | \ a) \textrm{ and } (x \ | \ b) \Rightarrow (x \ | \  \lambda a + \mu b) $.
\\
($ x \ | \ a  $ means "x divides a")
\\\\
For any $ x \in \Z_{\geq 0} $, $ (x \ | \ a) \textrm{ and } (x \ | \ b) \Rightarrow (x \ | \ \gcd(a, b)) $.
\\\\
If $  (x \ | \ ab) $ but $ \gcd(a, x) = 1 $ then $ (x \ | \ b ) $.
\\
(use Bezout's Theorem on $ \gcd(a, x) = 1 $)
\\
\begin{tcolorbox}[breakable]
Note: In general, $ \gcd(a, b) = 1 $ doesn't mean that $ (a \ \centernot| \ b) $, or vice versa.
\\\\
The only time that $ \gcd(a, b) = 1 \Leftrightarrow (a \ \centernot| \ b) $ is if one of $ a, b $ are prime. This is because (take $ p $ prime) the only way for $ \gcd(a, p) \neq 1 $ would be if $ a $ is a multiple of $ p $, in which case $ (p \ | \ a) $.
\end{tcolorbox}
\noindent
\\\\
\textit{Euclid's Algorithm}
\\
To find the $ gcd $ of $ a $ and $ b $ where $ a > b $. Write:
\begin{align}
a & = q_{1} b + r_{1} \\
b & = q_{2} r_{1} + r_{2} \\
r_{1} & = q_{3} r_{2} + r_{3} \\
& \vdots \\
r_{n-2} & = q_{n} r_{n-1} + r_{n}
\end{align}
Continue until $ r_{n} = 0 $, return $ r_{n-1} $.
\\
\begin{tcolorbox}[breakable]
In Haskell notation, Euclid's algorithm is specified by:
\begin{verbatim}
gcd(a, 0) = 0
gcd(a, b) = gcd(b, a mod b)
\end{verbatim}
\end{tcolorbox}
\noindent
\\
\textit{Bezout's Theorem}
\begin{equation} \label{bezout}
\gcd(a, b) = \lambda a + \mu b \quad (\textrm{for some } \lambda, \mu \in \R)
\end{equation}
(proof by induction)
\\
\begin{tcolorbox}[breakable]
We can write $ a $ and $ b $ as:
\begin{align}
a & = \alpha \gcd(a, b) \\
b & = \beta \gcd(a, b)
\end{align}
In general the solution to equation \ref{bezout} is given by:
\begin{equation}
\gcd(a, b) = (\lambda + \beta n) a + (\mu - \alpha n) b
\end{equation}
noting that the extra terms will always cancel out. So we have a set of solutions ($ \lambda_{n}, \mu_{n} $), where:
\begin{equation}
\lambda_{n} = \lambda + \beta n, \quad \mu_{n} = \mu - \alpha n
\end{equation}
\end{tcolorbox}
\noindent
\\
If a prime $ p \ | \ a_{1} a_{2} \textrm{...} a_{n} $ for $ a_{i} \in \Z $, then $ p $ divides any of the $ a_{i} $.
\\
(proof by induction on $ n $)
\\\\
Every integer larger than $ 1 $ can be written as a product of primes.
\\
(proof using strong induction)
\\\\
\textit{Fundamental Theorem of Arithmetic}
\\
Every integer larger than $ 1 $ can be written UNIQUELY as a product of primes.
\\
i.e. every such integer can be written as a product of primes $ p_{i}^{x_{i}} $, where each $ p_{i} $ is distinct and every $ x_{i} > 0 $.
\\
(proof not needed)

\subsection{Modular arithmetic}
$ (\equiv \mod m) $ is an equivalence relation.
\\\\
If $ a \equiv b \mod m $ and $ c \equiv d \mod m $:
\begin{align}
a + c & \equiv b + d \mod m \\
a - c & \equiv b - d \mod m \\
ac & \equiv bd \mod m
\end{align}
If $ a \equiv b \mod m $, then for any $ n $ :
\begin{equation}
a^{n} \equiv b^{n} \mod m
\end{equation}
(proof by induction)
\\\\
\textit{Fermat's Little Theorem}
\\
For any prime $ p $ and $ a \in \Z $:
\begin{equation}
a^{p} \equiv a \mod p
\end{equation}
(can be proven by induction)
\\\\
Also, if $ a $ is not a multiple of $ p $:
\begin{equation}
a^{p-1} \equiv 1 \mod p
\end{equation}
In the case $ a $ is a multiple of $ p $, we can easily see $ a^{p-1} \equiv 0 \mod p $.

\subsection{Equivalence relations and functions}
A binary relation $ \sim $ is an equivalence relation iff it is reflexive, symmetric and transitive.
\\\\
The equivalence class of $ b \in S $ is the set of all $ s \in S $ such that $ b \sim s $.
\\\\
The set of equivalence classes of a set $ S $ form a partition of $ S $. In other words, each equivalence class is a disjoint subset of $ S $.
\\\\
A function has an inverse iff it is a bijection.
\\\\
If $ f $ and $ g $ are injective, so is $ g \circ f $.
\\
If $ f $ and $ g $ are surjective, so is $ g \circ f $.
\\
If $ f $ and $ g $ are bijective, so is $ g \circ f $.
\\\\
If we define $ X \sim Y $ as "there exists a bijection $ f : X \to Y $", then $ \sim $ is an equivalence relation.
\\\\
Let $ P(A) $ denote the power set of a set $ A $.
\\
There exists no bijection $ A \to P(A) $.
\\\\
\textit{Examples of countably finite/infinite sets}
\\
$ \Z, \Q $ are countably infinite.
\\
$ \R, \C , \R \setminus \Q $ (set of all irrationals) are uncountably infinite.

\subsection{Combinatorics}
\textit{Multiplicative principle}
\\
In any $ n $-stage process, where the $ r^{th} $ stage can be completed in $ a_{r} $ ways, the total number of was of completing the process is:
\begin{equation}
\prod_{r = 0}^{n} a_{r}
\end{equation}
i.e. just multiply the number of options together.
\\\\
\\
\textit{Multinomials}
\\
When expanding $ (a + b + c + ...)^{n} $, with $ r $ terms inside the brackets, the general term of any coefficient is $ a^{\alpha} b^{\beta} c^{\gamma}... $, where $ \alpha + \beta + \gamma + ... = n $.
\\
The coeffient in the expansion for any term $ a^{\alpha} b^{\beta} c^{\gamma}... $ is given by:
\begin{equation}
\begin{pmatrix}
n \\ \alpha,  \beta,  \gamma ...
\end{pmatrix}
=
\frac{n!}{\alpha! \ \beta! \ \gamma! \ ... }
\end{equation}
\\
\begin{tcolorbox}[breakable]
Note if we set $ r = 2 $ in the above equation we get the familiar binomial coefficient:
\begin{align}
\begin{pmatrix}
n \\ \alpha,  \beta
\end{pmatrix}
& =
\frac{n!}{\alpha! \ \beta!} \\
& = \frac{n!}{\alpha! \ (n - \alpha)!}
\end{align}
since $ \beta = n - \alpha $.
\end{tcolorbox}

\end{document}